Paper 1: Triangulating trends in inequality around the world from secondary databases

### 2023-01-20 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
Adding to Paper 1 data analysis, shifting to python

---
title: "Appendix 1.A: Preparing PIP and WID inequality data"
format: html
warning: false
---

@@

#  A local version, in case internet isn't available 
df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")

@@

write.csv(df_pip, 
      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)

@@

# A local version, in case internet isn't available 
df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")

@@

write.csv(df_wid, 
      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)

@@

df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")

@@

df_regions<- df_regions %>%
  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
  rename(Entity = Our.World.In.Data.Name,
        WID_region = `region..standard.`) %>%
  mutate(region_alt = if_else(sub.region == 'Western Europe',
  'Western Europe',
  WID_region)) %>%
  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
  'Eastern Europe & Central Asia',
  WID_region)) %>%
  select(Entity, WID_region, region_alt)

kable(
  df_regions
)
```

Check that no countries are not matched for either dataset.
```{r}
# Merge in the regions dataset
df_pip<- left_join(df_pip, df_regions)
df_wid<- left_join(df_wid, df_regions)

# PIP check
kable(
  df_pip %>% 
    select(Entity, WID_region, region_alt) %>%
    unique() %>%
    filter(is.na(WID_region))
)

# WID check
kable(
  df_wid %>% 
    select(Entity, WID_region, region_alt) %>%
    unique() %>%
    filter(is.na(WID_region))
)

@@

Save the region mapping
```{r}
write.csv(df_regions, 
      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)

@@

@@

Save the population data
```{python}
# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')

>>>>

---
title: "Appendix 1.A: Preparing PIP and WID inequality data"
format: html
warning: false

---

@@

#  A local version, in case internet isn't available 
df_pip<- read.csv("data/original/pip_dataset.csv")

@@

write.csv(df_pip, 
      "data/clean/pip.csv", row.names=FALSE)

@@

# A local version, in case internet isn't available 
df_wid<- read.csv("data/original/wid.csv")

@@

write.csv(df_wid, 
      "data/clean/wid.csv", row.names=FALSE)

@@

# df_regions<- read.csv("data/original/WID_regions_country_standardized.csv")

```{python}

df_regions = pd.read_csv("data/original/WID_regions_country_standardized.csv")

@@

# df_regions<- df_regions %>%
#   select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
#   rename(Entity = Our.World.In.Data.Name,
#         WID_region = `region..standard.`) %>%
#   mutate(region_alt = if_else(sub.region == 'Western Europe',
#   'Western Europe',
#   WID_region)) %>%
#   mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
#   'Eastern Europe & Central Asia',
#   WID_region)) %>%
#   select(Entity, WID_region, region_alt)

# kable(
#   df_regions
# )
```

```{python}

df_regions = df_regions[['Our World In Data Name', 'region (standard)', 'sub-region']].rename(columns={'Our World In Data Name': 'Entity', 'region (standard)': 'WID_region'})

# Define an alt region classification
df_regions['region_alt'] = df_regions['WID_region']

df_regions.loc[df_regions['sub-region'] == 'Western Europe','region_alt'] = 'Western Europe'

df_regions.loc[(df_regions['sub-region'] == 'Eastern Europe') | (df_regions['WID_region'] == 'Russia & Central Asia'),'region_alt'] = 'Eastern Europe & Central Asia'

@@

Check that no countries are not matched for either dataset.
```{r}
# # Merge in the regions dataset
# df_pip<- left_join(df_pip, df_regions)
# df_wid<- left_join(df_wid, df_regions)

# # PIP check
# kable(
#   df_pip %>% 
#     select(Entity, WID_region, region_alt) %>%
#     unique() %>%
#     filter(is.na(WID_region))
# )

# # WID check
# kable(
#   df_wid %>% 
#     select(Entity, WID_region, region_alt) %>%
#     unique() %>%
#     filter(is.na(WID_region))
# )

@@

Save the region mapping
```{r}
# write.csv(df_regions, 
#       "data/clean/WID_region_mapping.csv", row.names=FALSE)

```{python}

df_regions.to_csv('data/clean/WID_region_mapping.csv')

@@

Check we have population for each country
```{python}
# df_pop = pd.read_csv('data/clean/population.csv')

# df_wid = pd.read_csv("data/clean/wid.csv")
# df_pip = pd.read_csv("data/clean/pip.csv")

# check_pip = pd.merge(df_pip, df_pop, how = 'left')
# check_pip[check_pip['population'].isna()]

# check_wid = pd.merge(df_wid, df_pop, how = 'left')
# check_wid[check_wid['population'].isna()]

# # There's no population for Zanzibar in the OWID data
# # For now I make up a population series based on two data points from Wikipedia. At some point I can come back and replaces this with the WID population data

# zanz_pop_1978 = 476111
# zanz_pop_2022 = 1889773

# # Assume a constant growth rate
# zanz_g = pow(zanz_pop_2022/zanz_pop_1978, 1/(2022-1978))-1

# year = 2022
# pop = zanz_pop_1978 * pow(1+zanz_g, year - 1978)

# df_zanz_pop = df_wid[df_wid['Entity'] == "Zanzibar"].copy()
# df_zanz_pop['population'] = zanz_pop_1978 * pow(1+zanz_g, df_zanz_pop['Year'] - 1978)

# df_zanz_pop = df_zanz_pop[['Entity', 'Year', 'population']]

# df_pop = pd.concat([df_pop, df_zanz_pop])

Save the population data
```{python}
# df_pop.to_csv('data/clean/population.csv')

### 2023-01-20 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Adding to Paper 1 data analysis, shifting to python

---
title: "Appendix 1.B: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive
---
```{python}

@@

df_wid = pd.read_csv("data/clean/wid.csv")
df_pip = pd.read_csv("data/clean/pip.csv")
df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")

@@

# df_wid = pd.read_csv("PhD_pages/data_appendices/data/clean/wid.csv")
# df_pip = pd.read_csv("PhD_pages/data_appendices/data/clean/pip.csv")
# df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")

@@

@@

@@

  plot = (ggplot(plot_data
  , aes(x_axis, y_axis))
  + geom_point(aes(color=region_col))
  + facet_wrap('~ source')
  + geom_polygon(aes(x='x', y='y'), data=shaded_coord, fill="#FF000066"))

@@

  # Count by rise/stable/fall bins

@@

  # Define the thresholds using the tolerance specified
  bin_thresholds = [-float("inf"), -tolerance, tolerance, float("inf")]
  bin_labels = ['fall', 'stable', 'rise']

  bins = pd.cut(plot_data['change'], bins=bin_thresholds, labels=bin_labels)

  # By region:

  # Count observations in bins defined by the thresholds
  threshold_counts = plot_data\
    .groupby(['source', region_col, bins])\
    .size()\
    .unstack()\
    .unstack(level=region_col)\
    .T\
    .reset_index()\
    .rename(columns={"change": "summary"})

  # Calculate average change and n
  mean_and_count = pd.DataFrame(plot_data.groupby(['source', region_col])['change']\
    .agg(['mean', 'count']))\
    .unstack()\
    .T\
    .reset_index()\
    .rename(columns={"level_0": "summary"})

  # rename the labels coming out of the aggregation
  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])

  # Concat together and present in a table
  df_summary_region = pd.concat([threshold_counts, mean_and_count])

  df_summary_region['summary'] = pd.Categorical(df_summary_region['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)

  df_summary_region = df_summary_region.sort_values([region_col,'summary'])

  # Set the index
  df_summary_region = df_summary_region.set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 5
  for i in range(0, len(df_summary_region.index.levels[0].unique())):
    df_summary_region.iloc[i*repeat_length,:] = df_summary_region.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[1 + i*repeat_length,:] = df_summary_region.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[2 + i*repeat_length,:] = df_summary_region.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[3 + i*repeat_length,:] = df_summary_region.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary_region.iloc[4 + i*repeat_length,:] = df_summary_region.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)

  # World summary

  # Count observations in bins defined by the thresholds
  threshold_counts = plot_data\
    .groupby(['source', bins])\
    .size()\
    .unstack(level='source')\
    .reset_index()\
    .rename(columns={"change": "summary"})

  # Calculate average change and n
  mean_and_count = pd.DataFrame(plot_data.groupby(['source'])['change']\
    .agg(['mean', 'count']))\
    .T\
    .reset_index()\
    .rename(columns={"index": "summary"})

  # rename the labels coming out of the aggregation
  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])

  # Concat together and present in a table
  df_summary_world = pd.concat([threshold_counts, mean_and_count])

  df_summary_world['summary'] = pd.Categorical(df_summary_world['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)

  df_summary_world = df_summary_world.sort_values(['summary']).set_index('summary')

  # format to show appropriate d.p. I do it row by row for want of figuring out a better way!
  df_summary_world.iloc[0,:] = df_summary_world.iloc[0,:].map('{:.0f}'.format)
  df_summary_world.iloc[1,:] = df_summary_world.iloc[1,:].map('{:.0f}'.format)
  df_summary_world.iloc[2,:] = df_summary_world.iloc[2,:].map('{:.0f}'.format)
  df_summary_world.iloc[3,:] = df_summary_world.iloc[3,:].map('{:.0f}'.format)
  df_summary_world.iloc[4,:] = df_summary_world.iloc[4,:].map('{:.2f}'.format)

  # ----- Return outputs ------
  plot_and_tables = {
    "table_data": plot_data,
    "table": plot_data_table,
    "plot": plot,
    "summary_world": df_summary_world,
    "summary_region": df_summary_region
  }

@@

# Specifications
reference_vals = [1990, 2018]
max_dist_from_refs = [4, 4]
min_dist_between = 20
value_col = 'Bottom_50_share'
# 'Gini'
# 'Top_10_share'
# 'Bottom_50_share'
# 'MLD'
# 'P90_P10_ratio' – not that helpful due to accuracy of low incomes

@@

@@

if value_col == 'Gini':
  tolerance = 0.02
  outlier_cut_off_upper = None

if value_col == 'Top_10_share':
  tolerance = 1
  outlier_cut_off_upper = None

if value_col == 'Bottom_50_share':
  tolerance = 1
  outlier_cut_off_upper = None

if value_col == 'P90_P10_ratio':
  tolerance = 0.5
  outlier_cut_off_upper = 40

@@

```{python}
plot_and_tables['plot']
```

```{python}
plot_and_tables['summary']
```

:::{.column-page}
```{python}

plot_and_tables['table'].show()

plot_data_results = plot_and_tables['table_data']

```
:::

>>>>

---
title: "Appendix 1.B: Analysing global trends in within-country inequality"
format: html
warning: false
filters:
  - shinylive

---
```{python}

@@

df_wid = pd.read_csv("data/clean/wid.csv")
df_pip = pd.read_csv("data/clean/pip.csv")
df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")
df_pop = pd.read_csv("data/clean/population.csv")

@@

# df_wid = pd.read_csv("PhD_pages/data_appendices/data/clean/wid.csv")
# df_pip = pd.read_csv("PhD_pages/data_appendices/data/clean/pip.csv")
# df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")
# df_pop = pd.read_csv("PhD_pages/data_appendices/data/clean/population.csv")

@@

  x_ref_val = f'{reference_col}{reference_vals[0]}'
  y_ref_val = f'{reference_col}{reference_vals[1]}'

@@

  # Add in population data
    # For first ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref1'})

    # For second ref
  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]

  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')

  plot_data = plot_data.rename(columns={'population':'population_ref2'})

    # Calculate average population two reference periods
  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2

@@

  plot = (ggplot(plot_data
  , aes(x_axis, y_axis, alpha=0.5))
  + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
  + facet_wrap('~ source')
  + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
  + theme_light())

@@

@@

  group_yes_vars = ['source', region_col]
  group_no_vars = ['source']

  group_scenarios = [group_no_vars, group_yes_vars]

  aggs = []

  for group_vars in group_scenarios:

  # Prepare pop-weighted average change
    agg_level = plot_data.copy()

    # Calulate pop weights by source and region
    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())

    agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())

    # Multiply the change by the pop weights 
    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']

  # Prepare fall/stable/rise categories
    agg_level['fall'] = agg_level['change'] < -tolerance

    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)

    agg_level['rise'] = agg_level['change'] > tolerance

    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)

    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
      .agg(['sum', 'mean', 'count'])

  # Aggregate by source and region
    if group_vars == group_yes_vars:

      agg_level = agg_level\
        .unstack().T.reset_index()

    elif group_vars == group_no_vars:

      agg_level = agg_level\
        .T.reset_index()

      agg_level[region_col] = "World"

    aggs.append(agg_level)

  df_summary = pd.concat(aggs)

    # filter for the aggregations we need and label in the 'summary' column
  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'

  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'

  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'

  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'

  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'

  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'

  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'

  df_summary = df_summary[df_summary['summary'].notnull()]

  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)

  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)

  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])

  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
  repeat_length = 7
  for i in range(0, len(df_summary.index.levels[0].unique())):
    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)

  # ----- Return outputs ------
  plot_and_tables = {
    "table_data": plot_data,
    "table": plot_data_table,
    "plot": plot,
    "summary": df_summary
  }

@@

```{python}
pd.set_option('display.max_rows', None)

## 1990-2017
```{python}
# Specifications
reference_vals = [1990, 2017]
max_dist_from_refs = [4, 4]
min_dist_between = 23

::: {.panel-tabset}
## Gini

```{python}
# | echo: false
# Specifications

value_col = 'Gini'
# 'Gini'
# 'Top_10_share'
# 'Bottom_50_share'
# 'MLD'
# 'P90_P10_ratio' – not that helpful due to accuracy of low incomes

tolerance = 0.02
outlier_cut_off_upper = None

```
::: {.panel-tabset}
## All countries
```{python}
# | echo: false

# Data from both sources is needed?
  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
  # If = 2, then only countries with data for both sources will be included
source_count_requirement = 1

```
::: {.panel-tabset}
## Plot
```{python}
plot_and_tables['plot']
```

## Summary
```{python}
plot_and_tables['summary']
```

## Data
```{python}

plot_data_results = plot_and_tables['table_data']
plot_data_results
```
:::

## Countries with data for both sources

```{python}
# | echo: false

@@

```
::: {.panel-tabset}
## Plot
```{python}
plot_and_tables['plot']
```

## Summary
```{python}
plot_and_tables['summary']
```

## Data
```{python}

plot_data_results = plot_and_tables['table_data']
plot_data_results
```
:::
:::

## Top 10% share

```{python}
# | echo: false
# Specifications

value_col = 'Top_10_share'

@@

tolerance = 1
outlier_cut_off_upper = None

```
::: {.panel-tabset}
## All countries
```{python}
# | echo: false

# Data from both sources is needed?
  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
  # If = 2, then only countries with data for both sources will be included
source_count_requirement = 1

@@

```
::: {.panel-tabset}
## Plot
```{python}
plot_and_tables['plot']
```

## Summary
```{python}
plot_and_tables['summary']
```

## Data
```{python}

plot_data_results = plot_and_tables['table_data']
plot_data_results
```
:::

## Countries with data for both sources

```{python}
# | echo: false

```
::: {.panel-tabset}
## Plot
```{python}
plot_and_tables['plot']
```

## Summary
```{python}
plot_and_tables['summary']
```

## Data
```{python}

plot_data_results = plot_and_tables['table_data']
plot_data_results
```
:::

:::

Next measure

:::

### 2023-01-20 PhD_pages/paper_pages/paper1.qmd
Adding to Paper 1 data analysis, shifting to python

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
---

```{r, echo=FALSE, message = FALSE, warning=FALSE}

library(tidyverse)
library(trackdown)

```

@@

### Other sources

Should we use Solt? As another check on results?

What about GCIP? For trying to flip between income and consumption...

@@

s

## Methods: Comparing databases

### Lining up years

Method 1: Reference years

Method 2: Fully extrapolating/interpolating

The equivalent of the reference year approach where there is no max threshold gap.

Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.

### Comparing welfare concepts

Getting a *global* view of inequalities within countries faces a Catch-22 situation: If you want broad coverage you get very heterogeneous data; if you want comparable data, or to be able to consider different welfare concepts consistently, the coverage you get is much smaller.

Here I'll take two avenues.

Firstly, what we're going for with the 'triangulation' approach is dominance - if different welfare metrics agree

Second we will cross reference to attempts to 'standardise' across these differences: using Solt and GCIP. How valid that is I don't know. But taking the two approaches together we get as close as we can to robust observations.

### A range of inequality metrics

### Lorenz curves

Dominance implies any (well-behaved) summary measure of inequality will agree on the ranking.

Operationalize this by taking 99 points on the Lorenze curve: the percentiles.

This leaves out the *very* top. But this would be quite a warped conception of inequality. I.e. if you have two distributions, A and B, and the share of income received by every whole percentile threshold in B is higher than A, but the Lorenze curves crossed within the top percentile - to claim that inequality is higher in B you'd need a very extreme conception of inequality.

# Results

### Outliers

Russia

>>>>

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
format: html
---

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages

```{r}
library(tidyverse)
library(knitr)
library(data.table)

```
:::

@@

@@

Give an overview of coverage in each dataset.

## Comparing databases

### Comparing metrics

Does it make a difference whether we look at Gini or top 10%?

People worry about whether Gini masks what's happening at the top.

Empirically: 
    - Make a linear model of top 10% on Gini, for each source separately
    - Make predicted top 10%
    - Take a look at the R2
    - Investigate individual countries with very poor predictions

Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.

### Since 1990

Similarities:

Convergence – Smaller rises in coutries that started out with high inequality. Driven regionally. (We could do a test for this.)

Indeed for some regions very similar picture:

    - MENA - Roughly 2 point fall, both Gini and top 10%. Only one country with a rise: Not total agreeement on coverage (one country either side)

    - South and South East Asia – Very similar not weighting by population. But wildly different when weighting by population, due to the much much stronger rise

    -  Sub-Saharan Afric – Very similar across sources. Top 10: Roughly even mix of rises and falling countries, but with a 3 point fall on average. Gini: More falls, 5 point on average. (Mention data quality issues: e.g. Kenya big fall. Lesoto big fall. But how comparable are the surveys?? Same surveys underlie WID...)

    - Western Europe – Rises in a majority of countries, but relatively moderate of 1-2 points (Whether Gini of top 10).

Differences:

Verry different bottom line at the global level:

- PIP: roughly equal numbers falling and rising. 0 average change – though a modest rise when pop weighted (2 points). Same whether Gini or Top 10. That's without LAtin America (Which WID is missing). Includung LatAm tips nudges it (with about a 1 point fall unweighter and 1 point rise weighted)

- WID: Far more countries with rising inequality. On average a 3 point rise (whether top 10 or Gini) – [Better to express these relative to the level of ineq. i.e. % increase rather than % point increase? A point rise means 'less' in WID terms because inequality levels are much higher.] But the most stark difference comes when we weight by population. Some key populous countries: India, China, US?, Indonesia? Russia

That difference is driven by particular regions:
-

More countries with rises. LArger rises. Especially larger populations. Average view very different:

Countries where the sources paint a radically different picture that affect the aggregate results:

- India: Insane rise in WID 34->56
- China:
- Russia: Massive drop in PIP. Insane rise in WID.
- USA: Much larger rise in WID.

What explains the different trends in these countries? (We can maybe look in more detail, at different welfare concepts etc.)

It sometimes goes the other way: Indonesia – Rise in PIP, fall in WID.

:::

#### Method 2: Use a regression to obtain the time trend

### Results for a range of inequality metrics

::: {.callout-note collapse="true" appearance="minimal"} 
## Calculations for comparisons

## Explaining differences between PIP and WID data

### For a subset of countries we can bridge between the PIP and WID data

Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.

In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:

-   WID has post-tax income series for some countries
-   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things. (For equivalisation, we could also possibly use the newly available PIP harmonized microdata).

### More in-depth discussion of particular outliers

-   Russia is a very notable case where the two datasets tell very different stories.
